\section*{Acknowledgments}

% To MIT
To my advisor Angela M. Belcher, and my thesis committee members K. Dane Wittrup and Cathy L. Drennan. They have supported as well as pushed my growth as a scientist and critical thinker. To the members of the Belcher Lab, who have provided me a home to pursue my research ambitions: Peter Jansen, Ngozi Eze, Archana Iyer, Dahyun Oh, John Casey, Matthew Klug, Maraym Moradi, Xiangnan Dang, Hiroshi Atsumi, Briana Dunn, Geran Zhang, Griffin Clausen, Nimrod Heldman, Jifa Qi, Ching-Wei Lin, Swati Kataria, Eric Lehnhardt, William Records, Tyler Toth, Uyanga Tsedev, Jelle Van Der Hilst, and Erin Reynolds. Special thanks to members of the Amon Lab who have patiently taught me the fundamentals of yeast biology, and made me appreciate the rigor and meticulousness that goes into the work of a pure biologist: Christopher Brennan, Summer Morrill, Daniel Corbi, Hilla Weidberg, and Xiaoxue ``Snow'' Zhou. Also, to the numberous scientists in the MIT scientific community who have helped, taught, and expedited my research with their expertise; they come from: the Center for Material Science \& Engineering, the Koch Facilities, the Whitehead Keck Microscopy Facility, the Center for Environmental Health \& Sciences Facility, the Department of Chemistry Instrumentation Facility, the Center for Bits and Atoms Media Lab, and the numerous maker spaces around MIT.

% To UC, Berkeley
To my previous advisors, Mikhail Shapiro and David Schaffer. My academic career at the University of California, Berkeley was arguably unparalleled and helped propel me to where I am today. I only wish I could be at the same intellectual and professional calibur as they.

% To Comm Lab
To the Communication Laboratory, Diana Chen, Prerna Bhargava, and Sean Clarke who have facilitated my growth as a speaker and educator. To David Ni\~no from the Gordon Engineering Leadership Program for his course 6.928J Leading Creative Teams which I helped taught. To Noreen Lyell, Leslie McClain, and Maxine Jonas for teaching an ambitious course, 20.109 Laboratory Fundamentals in Biological Engineering, and allowing me to teach alongside them.

% To Anthony Santella
To Anthony Santella, Professor at the University of New England, who enthusiastically introduced me to his scientific maker space, and including me in his student out-reach programs. I admire his ambition for promoting STEM outreach, and I was happy to have been involved.

% To the BE class
To my adopted academic family, the Biological Engineering PhD class: Ally Huang, Jared Kehe, Aaron and Mary Dy, Brett Geiger, Julia Joung, Manu Kumar, Andee Wallace, Ross and Katy Mitchell-Jones, Allison Tisdale, Erika Handly, Zijay Tang, Linyi Gao, Claire Duvallet, Malvika Verma, Paul Reginato, and everyone else from the program.

% Swing dancing
To the Boston swing dancing community: Dennis Walsh, Chris Warren, Katy Huang and Ray Zavesky, Kipaya Kapiga, Laura Nakolan, Elisabeth Sepulveda, and Surhbi Dhar. I could not have asked for better moments of laughter, conversation, and fun.

% California, friends
To my old friends from California: Raymong Yang and Lin Lin Phan, James Bui, Nathan Lin, Eric Mao, Felix Huang, Lauren Tang and Brian Paz, Lawrence Huang and Kristin Au, Julie Liang, Juna Song, Helen Sun, Selena Shen, Kunal Mehta, Jess Wang, and Jason Chavez. I hope to be back in the Bay Area, to relive our fun memories and hopefully make more.

% Family
What has driven me this far was my family. To my sister, Rosy, who has always been an example of empathy and kindness. To my brother-in-law, Lawrence, whose willingness to help others inspires me. To their daughters, my neices, Eleanor and Amelia Ong, who I hope I can make a better world for. To my in-laws, Grace and Eddie Ong, who have continuously supported our family. Finally, to my Mom and Dad; it has been my mission to prove to them that the sacrifices they have made, and are still making, will culminate to the success and enjoyment of mine and the future generations after me. I have truly stood on the shoulder of giants, and I hope not to squander the privileges I have been given.

\vspace{1.5\baselineskip}
\begin{center}
  \textbf{Thank you}
\end{center}
